---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
require(svglite)
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    dev = 'svglite',
    fig.path = "man/figures/README-",
    fig.showtext = TRUE,
    fig.width = 7,
    fig.height = 4,
    out.width = "100%",
    message = FALSE
)
```

# bondr

<!-- badges: start -->
[![R build status](https://github.com/dbarrows/bondr/workflows/R-CMD-check/badge.svg)](https://github.com/dbarrows/bondr/actions)
<!-- badges: end -->

Provides utilities and classes for working with reaction networks in R.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

```r
devtools::install_github("dbarrows/bondr")
```
## Creating networks

You write systems of reactions using a natural syntax, and bondr will parse it and turn it into an S3 object.

```{r}
library(bondr)

(synthesis <- parse_network("A + B -> C, 2.4e-5"))
```

### Sources / sinks

You can specify sources / sinks using `0` as the species name.

```{r}
parse_network("0 -> A, 4")
```

### Multiple reactions

Reactions can be entered on new lines.

```{r}
network_string <- "
    S + E -> SE, 1.66e-3
    SE -> E + P, 1e-1
"
parse_network(network_string)
```

### Bidirectional reactions

You can use `<->` to indicate bidirectional reactions, with an additional rate specified at the end of the line.

```{r}
parse_network("A <-> B, 1e-1, 2.2")
```

### Species orders

Prefixing a species name with a number will be interpreted as a reaction coefficient.

```{r}
network <- parse_network("2A -> B, 1e-1")
str(network$reactions[[1]])
```

## Using networks

### Propensity functions

You can generate the propensity functions for a reaction network.

```{r}
network <- parse_network("
         A -> B, 2.5
    2B + C -> A, 4e-2
")
(props <- propensities(network))
```

Note that dimerisations and multiple reactants are handled properly.

Propensity functions take a state vector of species quantities, ordered according to the output of the `species` function.

```{r}
species(network)
state <- c(2, 5, 4) # Corresponds to A, B, C
(props[[1]](state))
(props[[2]](state))
```

### Stoichiometric matrix

A matrix that conveys how the system updates when reactions fire. The columns correspond to reactions, and the rows to species.

```{r}
mm_string <- network_string_example("mm")
cat(mm_string)
(network <- parse_network(mm_string))
stoichiometric_matrix(network)
```

## Solving

`bondr` also provides a function `deriv_function` to get a derivative function compatible with the `deSolve` R package, which contains a number of numerical integrators.

### Using `deSolve`

Obtaining a deterministic solution to a system as in the Reaction Rate Equation can be done as follows.

```{r}
library(deSolve)

y <- c(S = 300, E = 120, SE = 0, P = 0)
times <- seq(0, 30, length.out = 100)
func <- deriv_function(network)

sol <- ode(y, times, func)
head(sol)
```

### Plotting

You can then plot the solution using a few `tidyverse` functions fairly easily.

```{r}
library(tidyverse)
library(mplot)

sol %>%
    data.frame() %>%
    rename(Time = time) %>%
    pivot_longer(species(network), names_to = "Species", values_to = "Quantity") %>%
    ggplot(aes(x = Time, y = Quantity, colour = Species)) +
        geom_line() +
        theme_mc()
```
