---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
require(svglite)
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = '#>',
    dev = 'svglite',
    fig.path = 'man/figures/README-',
    fig.showtext = TRUE,
    fig.width = 7,
    fig.height = 4,
    out.width = '100%',
    message = FALSE
)
```

# bondr

<!-- badges: start -->
[![R build status](https://github.com/dbarrows/bondr/workflows/R-CMD-check/badge.svg)](https://github.com/dbarrows/bondr/actions)
[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
<!-- badges: end -->

Provides utilities and classes for working with reaction networks in R.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

```r
devtools::install_github('dbarrows/bondr')
```
## Creating networks

You write systems of reactions using a natural syntax, and `bondr` will parse it and turn it into an S3 object.

```{r, include = FALSE}
devtools::load_all()
```
```{r, eval = FALSE}
library(bondr)
```
```{r}
(synthesis <- network('A + B -> C, 2.4e-5'))
```

### Sources / sinks

You can specify sources / sinks using `0` as the species name.

```{r}
network('0 -> A, 4')
```

### Multiple reactions

Additional reactions can be entered on new lines.

```{r}
network('
    S + E -> SE, 1.66e-3
    SE -> E + P, 1e-1
')
```

### Bidirectional reactions

You can use `<->` to indicate bidirectional reactions, with an additional rate specified at the end of the line.

```{r}
network('A <-> B, 1e-1, 2.2')
```

### Species orders

Prefixing a species name with a number will be interpreted as a reaction coefficient.

```{r}
net <- network('2A -> B, 1e-1')
str(net$reactions[[1]])
```

## Using networks

### Propensity functions

You can generate the propensity functions for a reaction network.

```{r}
net <- network('
         A -> B, 2.5
    2B + C -> A, 4e-2
')
(props <- propensities(net))
```

Note that dimerisations and multiple reactants are handled properly.

Propensity functions take a state vector of species quantities, ordered according to the output of the `species` function.

```{r}
species(net)
state <- c(2, 5, 4) # Corresponds to A, B, C
(props[[1]](state))
(props[[2]](state))
```

### Stoichiometric matrix

A matrix that conveys how the system updates when reactions fire. The columns correspond to reactions, and the rows to species.

```{r}
mm_string <- network_string_examples('mm')
cat(mm_string)
(net <- network(mm_string))
stmat(net)
```

## Solving

`bondr` also provides a function `deriv` to get a derivative function compatible with the [`deSolve`](https://cran.r-project.org/web/packages/deSolve/index.html) R package, which contains a number of numerical integrators.

### Using `deSolve`

Obtaining a deterministic solution to a system as in the Reaction Rate Equation can be done as follows.

```{r}
library(deSolve)

y <- c(S = 300, E = 120, SE = 0, P = 0)
times <- seq(0, 30, length.out = 100)
func <- deriv(net)

sol <- ode(y, times, func)
head(sol)
```

### Plotting

You can then plot the solution using a few [`tidyverse`](https://www.tidyverse.org/packages/) functions fairly easily.

```{r}
library(tidyverse)
library(wplot)

sol %>%
    data.frame() %>%
    rename(Time = time) %>%
    pivot_longer(species(net), names_to = 'Species', values_to = 'Quantity') %>%
    ggplot(aes(x = Time, y = Quantity, colour = Species)) +
        geom_line() +
        theme_wc()
```
