---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# chemnet

<!-- badges: start -->
<!-- badges: end -->

Chemnet provides utilities and classes for working with reaction networks

## Installation

You can install the development version from [GitHub](https://github.com/) with:

```r
devtools::install_github("dbarrows/chemnet")
```
## Creating networks

You write systems of reactions using a natural syntax, and chemnet will parse it and turn it into an S3 object.

```{r}
library(chemnet)

synthesis <- parse_network("A + B -> C, 2.4e-5")
synthesis
```

#### Sources / Sinks

You can specify sources / sinks using `0` as the species name.

```{r}
parse_network("0 -> A, 4")
```

#### Multiple reactions

Reactions can be entered on new lines.

```{r}
network_string <- "
    S + E -> SE, 1.66e-3
    SE -> E + P, 1e-1
"
parse_network(network_string)
```

#### Bidirectional reactions

You can use `<->` to indicate bidirectional reactions, with an additional rate specified at the end of the line.

```{r}
parse_network("A <-> B, 1e-1, 2.2")
```

#### Species Orders

Prefixing a species name with a number will be interpreted as a reaction coefficient.

```{r}
network <- parse_network("2A -> B, 1e-1")
str(network$reactions[[1]])
```

## Using networks

#### Propensity functions

You can generate the propensity functions for a reaction network.

```{r}
network <- parse_network("
         A -> B, 2.5
    2B + C -> A, 4e-2
")
props <- propensities(network)
props
```

Note that dimerisations and multiple reactants are handled properly.

Propensity functions take a state vector of species quantities, ordered according to the output of the `species` function.

```{r}
species(network)
state <- c(2, 5, 4) # Corresponds to A, B, C
(props[[1]](state))
(props[[2]](state))
```
